Limited physiological plasticity and reduced relative variance in physiological rates of ectotherm populations under climate change

### 2024-07-09 ms.qmd
Add Sinclair ref which is important

We show that acclimation of physiological rates and changes in relative
variance differ across habitats. Our findings confirm previous results
that quantify the different capacity of terrestrial, marine and
freshwater ectotherms to acclimate [@Seebacher2015-wi;
@Gunderson2015-hz; @morley2019physiological]. Our analysis generally
supports that of @Seebacher2015-wi, @Gunderson2015-hz and
@morley2019physiological that all show a general inability of
terrestrial ectotherms to physiologically acclimate. The change in
acclimation $Q_{10}$ we found was similar to @Seebacher2015-wi for
freshwater organisms (~17%), but higher in marine ectotherms [decrease
of 16% versus ~10% in @Seebacher2015-wi], and opposite in terrestrial
ectotherms [increase of ~6% compared to an ~8% decrease in
@Seebacher2015-wi], possibly due to the use of newly derived
$Q_{10}$ effect sizes that allow us to control for sampling variance.
Greater capacity for acclimation in aquatic organisms may be the result
of fewer opportunities for behavioural thermoregulation in aquatic
environments making physiological remodeling important for maintaining
homeostasis [@Gunderson2015-hz; @morley2019physiological].
Additionally, aquatic and terrestrial environments can vary in
nutritional resources [e.g., fatty acid content of prey;
@twining2019aquatic] that can impact physiological rates and their
variability by affecting performance curves and acclimation responses
[@hardison2024diet]. While marine and freshwater ectotherms were
capable of partial acclimation it is still important to recognise that,
on average, the effect size was small (amounting to $Q_{10}$ dropping
from ~1.8 to 1.6).

@@

Acclimation capacities are expected to differ between life-stages
because of distinct of patterns of dispersal, habitat use and behaviour
that force earlier life stages to cope with more variable environmental
conditions which can also lead to developmental constraints on how
physiological systems respond later in life [@martin2015age;
@stearns1976life; @angilletta2009thermal; @noble2018developmental;
@o2019developmental; @Pottier2022]. In addition, plastic responses are
also expected to be costly [@angilletta2009thermal; @Dewitt1998], and
such costs can be magnified in late life reducing the capacity for
plasticity [e.g., @rossi2019ageing]. These processes can also result in
changes to intrapopulation variation in physiological rates at higher
temperatures but the direction of change between early and adult life
stages is likely to depend on the costs of adjusting physiological
processes, energy reserves at different life stages, and the extent to
which early life experiences constrain plasticity.

>>>>

We show that acclimation of physiological rates and changes in relative
variance differ across habitats. Our findings confirm previous results
that quantify the different capacity of terrestrial, marine and
freshwater ectotherms to acclimate [@Seebacher2015-wi;
@Gunderson2015-hz; @morley2019physiological]. Our analysis generally
supports that of @Seebacher2015-wi, @Gunderson2015-hz and
@morley2019physiological that all show a general inability of
terrestrial ectotherms to physiologically acclimate. The change in
acclimation $Q_{10}$ we found was similar to @Seebacher2015-wi for
freshwater organisms (~17%), but higher in marine ectotherms [decrease
of 16% versus ~10% in @Seebacher2015-wi], and opposite in terrestrial
ectotherms [increase of ~6% compared to an ~8% decrease in
@Seebacher2015-wi], possibly due to the use of newly derived
$Q_{10}$ effect sizes that allow us to control for sampling variance.
Greater capacity for acclimation in aquatic organisms may be the result
of fewer opportunities for behavioural thermoregulation in aquatic
environments making physiological remodeling important for maintaining
homeostasis [@Gunderson2015-hz; @morley2019physiological].
Additionally, aquatic and terrestrial environments can vary in
nutritional resources [e.g., fatty acid content of prey;
@twining2019aquatic] that can impact physiological rates and their
variability by affecting performance curves and acclimation responses
[@hardison2024diet; @sinclair2016can]. While marine and freshwater ectotherms were
capable of partial acclimation it is still important to recognise that,
on average, the effect size was small (amounting to $Q_{10}$ dropping
from ~1.8 to 1.6).

@@

Acclimation capacities are expected to differ between life-stages
because of distinct of patterns of dispersal, habitat use and behaviour
that force earlier life stages to cope with more variable environmental
conditions which can also lead to developmental constraints on how
physiological systems respond later in life [@martin2015age;
@stearns1976life; @angilletta2009thermal; @noble2018developmental;
@o2019developmental; @Pottier2022; @sinclair2016can]. In addition, plastic responses are
also expected to be costly [@angilletta2009thermal; @Dewitt1998], and
such costs can be magnified in late life reducing the capacity for
plasticity [e.g., @rossi2019ageing]. These processes can also result in
changes to intrapopulation variation in physiological rates at higher
temperatures but the direction of change between early and adult life
stages is likely to depend on the costs of adjusting physiological
processes, energy reserves at different life stages, and the extent to
which early life experiences constrain plasticity.
